---
title: "Media Mobilizing Action: Understanding the Media as a Signal for Bureaucratic Action 
Megan Martinez - Gov 51 Final Project"
format: pdf
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

Github: <https://github.com/meganmartinez-25/Gov-51-Final-Project/tree/main>

## How does media coverage impact bureaucratic action?

Introduction:

In 2014, massive surges in unaccompanied migrants illegally traveling to the US caught the attention of the media. Public opinion polls at the time found that one in six Americans listed immigration as the most important problem in the U.S. (Saad 2014). Gallup polls attributed this rise in concern about immigration to a rise in a greater “news focus on tens of thousands of refugees from Central American countries attempting to enter the U.S. at its southern border” (Newport 2018). 

While the media plays a central role in influencing the topic of most importance to the public, it is unclear if its coverage influences bureaucratic agents and the actions they take. Does media coverage of immigration serve as a direct signal that influences bureaucratic behavior, as measured by changes in Immigration and Customs Enforcement (ICE) arrest activity? 

In this paper, I seek to fill the gap of the role of media as a direct signal to bureaucrats, ultimately influencing their behavior through the content they hear and see rather than through public opinion pressures. Bureaucratic agencies seeking to maintain their presence and authority will likely be quick to respond to the media that signals the issues that their elected supervisors are also monitoring. I argue that because of this expectation, negative media coverage about immigration will be associated with an increase in ICE arrests, while positive or no sentiment media would be associated with a slight decrease or no change in ICE arrests. 

To explore the phenomenon I will review the relevant literature. I examine the media’s influence on agenda-setting, bureaucratic agencies’ decision making, and the potential relationship between the two. In my methods section, I propose a time series analysis to assess the visual relationship between media produced on immigration and ICE arrest rates over time. This study will be effective in examining if ICE arrests fluctuate based on the sentiment of the media on immigration produced (negative or positive). 

**Role of Media and Agenda-Setting**

The media has the capacity not to tell people what they ought to be thinking, but what they should be “thinking about” (Cohen 1963, 13). The content that editors choose to publish acts as an additional signal to the public that these topics are worth discussing. The theory of agenda-setting argues that mass media raises attention around certain stories and people to influence what individuals should “think about, know about, have feelings about” (Lang and Lang 1966, 468 cited in McCombs and Shaw 1972, 177). McCombs and Shaw tested this theory and found a 0.979 correlation between issues emphasized by the media during a campaign and what voters believed was important, accounting for issues that aligned with one’s party ideals (McCombs and Shaw 1972, 181-182).

          The agenda-setting nature of the media to influence public opinion is proportionate to the number of stories published around an issue.  In Iyengar and Kinder’s experiment on the variation of news coverage’s influence on one’s judgment of the president’s performance, their results validated their priming theory arguing that a greater volume of coverage on an issue means that issue will hold more weight in a president’s evaluation (Iyengar and Kinder 1987, 70). Following the persuasion bias theory, when one does not account for the repetition of information that they receive, over time their views may adapt to the type of “slanted” media they are consuming (DeMarzo et al. 2003, 1-2). Viewers are not only sourcing their information from the media, but being subconsciously influenced by what they repeatedly see, which can influence their beliefs and actions.                                                                         

**Bureaucracy and Literature**

Bureaucrats, as civil servants, do not face public expectations to maintain their positions in government. Instead, as appointed officials, they are often selected based on skill and expertise and derive their power “from above,” or from the political appointee, but have a degree of agency to make decisions to perform their role  (Weber 1946, 201). 

Yet, bureaucrats are granted civil service liberties that protect them from exploitation or manipulation, and grants the public a sense of security that agencies will continue to function regardless of the party in power (Huber and Ting 2015, 3). As a result, bureaucrats can act independently of social pressure, which allows them to mitigate the call by public opinion and potentially rely more on sources of media to indicate issues in communities that they do not have access to. 

Because bureaucrats have greater autonomy in their actions, they may be swayed by other indicators. For one, private interests often dominate the spectrum of public discussions as many government appointees previously served in the private sector (Domhoff et al. 2017, 25). The entrance model suggests that when private sector employees enter into the public sphere, they are more likely to take action that support corporations than those who did not work in the private sector, and may be ingrained with biases that can be considered more private sector leaning (Vandenbergh et al. 2020, 1130). 

Elites and interest groups may also influence bureaucrats more directly through lobbying rather than using media as a signal. In a study that analyzed 40 rules across a variety of bureaucratic agencies, when business comments dominated (about 90% of all comments), government increased only 4% of the time or there was no change to government involvement 10% of the time (Yackee and Yackee 2006, 131 & 136). Corporations may also sponsor grassroots organizations to lobby in order to gather a large public voice behind a certain issue on behalf of other agents (Walker 2009, 83). In fact, corporations make up 53% of the clients for grassroots organizations (Walker 2009, 91). As I consider the relationship between media and bureaucracy, this study will be setting up a contrast with the leverage elites’ hold over bureaucracies. 

**Immigration, Media, and Bureaucracy**

       Previous sections demonstrate that the media sets the agenda for public officials. Literature on bureaucracy indicates that people who are supposed to be apolitical can be influenced by their surrounding environment. However, there is a gap in understanding whether media narratives directly influence agencies, and how. 

       One possible explanation is that the media can create social imaginaries that become popularized rhetorics used against outsider groups to influence dialogue around certain policy questions. For example, the media can create social imaginaries that weaponize subjectivities about identities to create narratives that turn stories into truth, as seen through the rise of the “Latino Threat Narrative” that likens Latin immigrants to aliens and invaders (Chavez 2013, 54). England faced a similar situation when waves of petty crimes from the youth led to an overdramaticization of the mods and rockers as “happy hoodlums” and “drug crazed youth” in the media, which eventually became a societal understanding rather than an inflated rhetoric (Cohen 2002, 151). While originally the Home Secretary failed to submit to this new rhetoric calling it an exaggeration, he ultimately supported the Malicious Damage Bill that wishes to punish these youth, in order to put himself on the right side of the battleline and protect his image (Cohen 2002, 153-155). Ultimately, this case demonstrates the same path that immigration is taking in the U.S. As stories surrounding immigration portray immigrants in a severe negative light, bureaucrats may be forced to choose a side, taking extreme measures by using policy to appear to take action to what the media is portraying as an “evil” in society. 

       Another explanation would claim that bureaucracy responds to media because media reflects the interest of elite actors, who drive their decision-making. Corporate capture is the phenomena of political decisions being made in the best interests of private industries at the expense of the public good (Bordonada 2022). Elite individuals may influence the type of media or stories that are produced to push policies they want. Some research finds that corporate media outlets tend to avoid stories that undermine the economic interests of their owners. (Gilens and Hertzman 2003, 370). Nevertheless, this linkage has not yet been tested and it would be difficult to track if corporate executives were specifically ordering journalists to increase their negative rhetoric in stories on immigration. Also, not every media organization is elite owned and it is possible that the effect of media coverage on bureaucratic action is independent of that organization’s owner. As the literature review has discussed, the media’s power on agenda-setting does not generally depend on the type of ownership model of the organization (including for profit, non-profit, and governmental). Instead, the content and repetition of the news seems to drive policy, among other things, and my research will seek to understand why policy may be reflective of trends in the media. 

       Ultimately, because the media serves as a source of information signaling to government agencies, there should be a significant positive relationship between negative national media coverage and a bureaucracy’s action. 

Expectation 1: Negative national media coverage is associated with an increase in bureaucratic policymaking. 

       On the other hand, if the story does not contradict the agency’s prerogative, agencies are unlikely to change their behaviors. Agencies have limited missions, and they will filter stories that contradict or do not relate to their goals. They will likely be satisfied with their progress and continue doing their work. 

Expectation 2: Positive or neutral national media coverage is associated with no increase or a slight decrease in bureaucratic policymaking. 

**Date**

To capture the the volume of media covered around immigration, I will be using TV News Channels, that holds 23 sources and contains “content from the closed captions of major american TV news stations and shows. Each source is the TV station; each story represents a "segment" within a show; the title of a story is the TV show the segment is from. This content, and the automated segment detection, is thanks to the MIT Media Lab SuperGlue project” ([MediaCloud.org](http://mediacloud.org/)). Through Media Cloud, I filtered to phrases I associated with immigration, including “ICE,” “Immigration,” and “Immigrants.” This would allow me to capture the volume of News segments that included these phrases per day, which i then sorted into weeks in order to match the data from Deportation Data. In addition, I selected to data from September 1st, 2023 through March 2026, in order to match the timeline for the data on arrests. This data set contains four columns, including the date, count of segments that made reference to immigration, the total number of segments on that day, and the proportion of immigration related segments out of total segments produced for the day.

To capture data on ICE enforcement, I will utilize the “ICE Administrative Arrests” Data gathered from the Deportation Data Project, an independent organization that discloses “U.S. government immigration enforcement data via public records litigation” (“Data Deportation Project”). The dataset contains all arrests made each day from September 1st, 2023 to March 10th, 2026, containing specific information about each individual detained, including reason for detention, area of responsibility their arrest falls under, detention state, date of detention, severity of criminality on a three point scale, and additional metrics. I will be focusing on the date of detention, the area of responsibility, and the state where the arrest occurred. During this time period, there are 713,464 arrests coded. I will be filtering my data to calculate the count of individuals arrested per week. 

**Methods**

To examine the relationship between media coverage and ICE enforcement activity, this paper begins by conducting a time series analysis, which is suitable to detect if shifts in media framing precede ICE enforcement over time. I will examine national media coverage of national ICE arrests to determine if there is a lagged relationship where media spikes precede spikes in ICE arrests by 3-6 weeks. 

In order to compare media coverage on immigration to the proportion of ICE arrests, I would use a time series plot. To determine if ICE arrests were a reaction to media coverage, I would expect to see a spike in the proportion of ICE arrests 3-6 weeks after spikes in media coverage on immigration, indicating that bureaucratic action was taken to increase enforcement, demonstrated in the proxy of arrests. This lag reflects the time needed for ICE bureaucrats to consistently observe media narratives and translate these signals into enforcement decisions, seen through the proxy of additional ICE arrests. If there is no relationship, I would first conduct robustness checks to ensure that my time lag is not influencing my results, and instead, examine if adapting the time period to be longer or shorter may more accurately reflect a significant relationship. 

If robustness checks across varying lag lengths produce no significant relationship, I would fail to reject the null hypothesis that media coverage of immigration has no statistically discernible relationship with ICE enforcement activity. Instead, media and enforcement spikes may be a result of chance or external events, such as greater number of border crossings, changes in federal policy, and shifts in public opinions, rather than a direct media signal to bureaucratic agencies. 

Part 1: Reading in the data

<https://github.com/meganmartinez-25/Gov-51-Final-Project>

```{r}
knitr:: opts_chunk$set (echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(readxl)
library(kableExtra)

arrests_raw <- read.csv("data/raw/arrests_state_filled.csv")
media_raw <- read.csv("data/raw/mc_TVcoverage.csv")
liberal_media <- read.csv("data/raw/mc-onlinenews-mediacloud-20260511025540-counts.csv")
conservative_media <- read.csv("data/raw/mc-onlinenews-mediacloud-20260511025846-counts.csv")


```

**Summary Statistics:**

```{r, results = 'asis'}

arrests_state <- arrests_raw |> 
  filter(!is.na(apprehension_state)) |> 
  count(apprehension_state, sort = TRUE) |> 
  mutate(prop = round(n/sum(n), 2)) |> 
  kable(
    format = "latex", 
    booktabs = TRUE,
    longtable = TRUE, 
    col.names = c("State", "Count", "Proportion"), 
    caption = "ICE Arrests by State, Oct. 2022- Mar. 2026 "
  ) |> 
  kable_styling(latex_options = c("repeat_header"))

arrests_threat <- arrests_raw |> 
  filter(!is.na(case_threat_level)) |> 
  count(case_threat_level, sort = TRUE) |> 
  mutate(prop = round(n/sum(n), 2)) |> 
  kable(
    format = "latex", 
    booktabs = TRUE, 
    col.names = c("Threat Level", "Count", "Proportion"), 
    caption = "ICE Arrests by Threat Level, Oct. 2022- Mar. 2026 "
  )

arrests_gender <- arrests_raw |> 
  filter(!is.na(gender), 
         gender != "", gender != "Unknown") |> 
  count(gender, sort = TRUE) |> 
  mutate(prop = round(n/sum(n), 2)) |> 
  kable(
    format = "latex", 
    booktabs = TRUE, 
    col.names = c("Gender", "Count", "Proportion"), 
    caption = "ICE Arrests by Gender, Oct. 2022- Mar. 2026"
  )

arrests_birth_year <- arrests_raw |> 
  filter(!is.na(birth_year)) |>
  mutate(arrest_age_group = case_when( 
    2026 - birth_year <18 ~ "Under 18", 
    2026 - birth_year >= 18 & 2026- birth_year <= 25 ~ "18-25", 
    2026 - birth_year >= 26 & 2026- birth_year <= 35 ~ "26-35",  
    2026 - birth_year >= 36 & 2026- birth_year <= 45 ~ "36-45", 
    2026 - birth_year >=46 ~ "Over 45")) |> 
  count(arrest_age_group, sort = TRUE) |> 
  mutate(prop = round(n/sum(n), 4)) |> 
  kable(
    format = "latex", 
    booktabs = TRUE, 
    col.names = c("Age Group", "Count", "Proportion"), 
    caption = "ICE Arrests by Birth Year, Oct. 2022- Mar. 2026 "
  )


#arrests_date
arrests_state
cat("\\clearpage")

arrests_threat

arrests_gender

arrests_birth_year
cat("\\clearpage")


```

       Due to majority of my critical variables being character strings, I chose to do counts and proportions for my summary statistics, to better understand the patterns, most common characters within each variable, and the distribution between the characters. This is crucial to understand because my analysis focuses on the volume and distribution over time, so with variables coded as words, it is important to have an accurate reading and representation of these.

       The first table describes the number of ICE arrests from Oct. 2022- Mar. 2026 in each state, and the proportion of ICE arrests in each state to the total nationwide number of arrests. This allowed me to see which states had the most ICE arrests during this time period, which is a signaler for places that may or may not have received high media coverage. The second table allowed me to do the same counts and proportions as the first, however, I am specifically looking at what ICE classifies as threat levels, and what proportion of arrested individuals were within each category. This process was conducted again for gender and age, to understand the distribution of what individuals are actually being arrested and at what age.

**ICE Arrests By State Map**

```{r}
library(usmap)

arrest_state_count <- arrests_raw |> 
   filter(!is.na(apprehension_state)) |> 
  count(apprehension_state, sort = TRUE) |> 
  mutate(prop = round(n/sum(n), 2)) |> 
  rename(state = apprehension_state)

plot_usmap(data= arrest_state_count, values = "n", color = "white") +
  scale_fill_continuous(
    low = "lightblue", high = "darkblue", 
    name = "ICE Arrests"
  ) + 
  labs( 
    title = "ICE Arrests by State, Oct. 2022- Mar. 2026")
```

This map specifically indicates through a color gradient, the prevalence of arrests in certain states. Specifically, I was able to recognize Texas, California, Florida and New York being high centers of ICE arrests, which would allow me to increase my focus on which regions I could use in future case studies. The darker the color of the state, the greater number of arrests located in that state proportional to the total number of ICE arrests from October 2022 to March 2026.

**Time Series of Counts of ICE Arrests Per Week**

```{r}
arrests_weekly <- arrests_raw |> 
  filter(!is.na(apprehension_date)) |>
  mutate(week = floor_date(as.Date(apprehension_date), "week")) |>
  count(week, name = "arrests") 
  
key_date <- data.frame(
  date = as.Date(c("2024-11-05", "2025-01-20")), 
  label = c("2024\nelection", "Trump\ninauguration"), 
  y_pos = c(9750, 8000))

arrests_line <- ggplot(arrests_weekly, aes( x= week, y = arrests)) + 
  geom_area(fill = "darkblue", alpha = .4) + 
  geom_line(color = "darkblue") + 
  geom_vline(data = key_date, aes(xintercept = date), 
             color = "red", linetype = "dashed") + 
  geom_text(data = key_date, aes(x = date, label = label, y = y_pos)) + 
  scale_x_date(limits = as.Date(c("2022-10-01", "2026-03-31"))) + 
  labs(title = "Ice Arrests Per Week from Oct. 2022 - Mar. 2026",
       x = "Oct. 2022 - Mar. 2026", 
                y = "Count of Arrests per Week")
   
arrests_line

```

The line graph looks to map weekly the number of ICE arrests over time from 2022-2026. Specifically, mapping the peaks of political events, such as the 2024 election and Trump's inauguration allows us to also see and predict the correlation of these events to the spikes we see in the data for ICE arrests. This could also indicate moments of high media volatility and could also demonstrate a potential linkage between media and ICE arrests. The graph begins to peak and rise after President Trump's inauguration on January 20th, 2025. However, there still are spikes around ICE arrests in 2023 to early 2024.

**Time Series for the Proportion of ICE Arrests Per Week**

```{r}
arrests_weekly_prop <- arrests_raw |> 
  filter(!is.na(apprehension_date)) |>
  mutate(week = floor_date(as.Date(apprehension_date), "week")) |>
  count(week, name = "arrests") |> 
  mutate(arrests_prop = arrests/ sum(arrests))
  
key_date <- data.frame(
  date = as.Date(c("2024-11-05", "2025-01-20")), 
  label = c("2024\nelection", "Trump\ninauguration"), 
  y_pos = c(0.012, .018))

arrests_line_prop <- ggplot(arrests_weekly_prop, aes( x= week, y = arrests_prop)) + 
  geom_area(fill = "darkblue", alpha = .4) + 
  geom_line(color = "darkblue") + 
  geom_vline(data = key_date, aes(xintercept = date), 
             color = "red", linetype = "dashed") + 
  geom_text(data = key_date, aes(x = date, label = label, y = y_pos)) + 
  scale_x_date(limits = as.Date(c("2022-10-01", "2026-03-31"))) + 
  scale_y_continuous(limits = c(0,.020)) +
  labs(title = "Ice Arrests Per Week from Oct. 2022 - Mar. 2026",
       x = "Oct. 2022 - Mar. 2026", 
                y = "Proportion of Arrests per Week")
   
arrests_line_prop
```

The graph above indicated the proportion of ICE arrests per week from October 2022 to March 2026. This allows us to see the same patters as above with the ICE Arrests per Week in counts, but by using proportions, we are better able to compare to the media that will also be looked at in proportions.

**Time Series for the Count of Media Segments Produced on Immigration per Week**

```{r}
media_weekly <- media_raw |> 
  filter(!is.na(date)) |>
  mutate(week = floor_date(as.Date(date), "week")) |>
  group_by(week) |> 
  summarize(media_count = sum(count, na.rm = TRUE), 
            media_total = sum(total_count, na.rm=TRUE)) 

key_date <- data.frame(
  date = as.Date(c("2024-11-05", "2025-01-20")), 
  label = c("2024\nelection", "Trump\ninauguration"), 
  y_pos = c(7000, 5000))
  

media_line <- ggplot(media_weekly, aes( x= week, y = media_count)) + 
  geom_area(fill = "darkred", alpha = .4) + 
  geom_line(color = "darkred") + 
  geom_vline(data = key_date, aes(xintercept = date), 
             color = "red", linetype = "dashed") + 
  geom_text(data = key_date, aes(x = date, label = label, y = y_pos)) + 
  scale_x_date(limits = as.Date(c("2022-10-01", "2026-03-31"))) + 
  labs(title = "Media Segments on Immigration Per Week from Oct. 2022 - Mar. 2026",
       x = "Oct. 2022 - Mar. 2026", 
                y = "Count of TV Segments related to immigration per Week")
   
media_line

```

The figure above allows us to see the count of media segments related to immigration per week from October 2022 to March 2026. However, because immigration is only a small portion of the total media covered per week, it can be difficult to see the change over time. Therefore, I decided to use a proportion graph to better see how the proportion of media referencing immigration changes per week.

**Time Series Analysis for the Proportion of Media Segments on Immigration Per Week**

```{r}

media_weekly <- media_raw |> 
  filter(!is.na(date)) |>
  mutate(week = floor_date(as.Date(date), "week")) |>
  group_by(week) |> 
  summarize(media_count = sum(count, na.rm = TRUE), 
            media_total = sum(total_count, na.rm=TRUE), 
            media_prop = media_count / media_total) 

key_date <- data.frame(
  date = as.Date(c("2024-11-05", "2025-01-20")), 
  label = c("2024\nelection", "Trump\ninauguration"), 
  y_pos = c(.19, .23))
  

media_line_prop <- ggplot(media_weekly, aes( x= week, y = media_prop)) + 
  geom_area(fill = "darkred", alpha = .4) + 
  geom_line(color = "darkred") + 
  geom_vline(data = key_date, aes(xintercept = date), 
             color = "red", linetype = "dashed") + 
  geom_text(data = key_date, aes(x = date, label = label, y = y_pos)) + 
  scale_x_date(limits = as.Date(c("2022-10-01", "2026-03-31"))) + 
  scale_y_continuous(limits = c(0,.25)) + 
  labs(title = "Media Coverage on Immigration from Oct. 2022 - Mar. 2026",
       x = "Oct. 2022 - Mar. 2026", 
                y = "Count of TV Segments related to immigration per Week")
   
media_line_prop
```

The graph above looks at the proportion of media segments per week that reference immigration out of the total media produced that week. This looks how the proportion of media on immigration shifted over time. We are able to see a recent spike in 2026, but also a spike mid-202, and leading up to 2025. There is a sharp increase on proportion of media discussing immigration around Trump's inauguration which may be an effect of his discussions of increasing immigration policy when entering office.

**Unify the Media and ICE Arrests Data Together**

```{r}

library(tidyverse)
combined <- arrests_weekly |> 
  inner_join(media_weekly, by = "week") |> 
    filter(week != as.Date("2022-09-25")) |> 
  arrange(week) |>
  mutate(media_pct = media_prop * 100) |> 
  mutate(
    log_arrests = log(arrests), 
    media_lag1 = lag(media_pct, 1), 
    media_lag2 = lag(media_pct, 2), 
    media_lag4 = lag(media_pct, 4),
    media_lag8 = lag(media_pct, 8)
  )


```

**Correlations**

```{r}
library(tidyverse)
no_week_cor <-cor.test(combined$media_pct, combined$log_arrests)

 week1_cor <-cor.test(combined$media_lag1, combined$log_arrests)
week2_cor <- cor.test(combined$media_lag2, combined$log_arrests)
week3_cor <- cor.test(combined$media_lag4, combined$log_arrests)
week4_cor <- cor.test(combined$media_lag8, combined$log_arrests)



correlations <- data.frame(
  lag = c("No Lag", "1-Week Lag", "2-Week Lag", "4-Week Lag", "8-Week Lag"),
  Correlation = round(c(
    no_week_cor$estimate, 
    week1_cor$estimate, 
    week2_cor$estimate, 
    week3_cor$estimate, 
    week4_cor$estimate
  ), 3), 
  P_value = round(c(
    no_week_cor$p.value, 
    week1_cor$p.value, 
    week2_cor$p.value, 
    week3_cor$p.value, 
    week4_cor$p.value
  ), 15)
) |> mutate(
  Significant = case_when(
    P_value < 0.001 ~ "***", 
      P_value < 0.01 ~ "**", 
      P_value < 0.05 ~ "*", 
      TRUE ~ ""
  )
) |> 
  kable(
    booktabs = TRUE, 
    col.names = c("Lag", "Correlation", "P-Value", "Significance"), 
    caption = "Correlation between media coverage on immigration and Log ICE arrests"
  ) 
  

correlations
```

All five of these correlations ( with time lags and without), have a moderate, positive correlation. This indicates that when there are weeks with higher media coverage, there is likely to be higher ICE arrests that week as well. The highest correlation is found when there is no time lag ( r = 0.486). As the lag increases, the correlation gets weaker, and by the 8th week after media coverage on immigration has risen, there is not as high of a rise in ICE arrests (r= 0.386). Therefore, this could indicate that over time, media moves alongside ICE arrests in similar patterns that are highly correlated in shorter time periods. Therefore, media may not be as effective as a predictor of ICE arrests as previously hypothesized. Because all correlations have a p-value less than 0.01, there is a very low probability that one would see this correlation if there was no relationship between media coverage on immigration and ICE arrests. The likelihood of finding this correlation by random chance is unlikely; therefore, there must be some relationship between the two variables.

However, because this is a correlation we cannot claim that there is a causal relationship between media coverage on immigration and the log of ICE Arrests. Instead, we can claim that we are confident that this relationship exists between the two variables.

**CCF Graph**

```{r}
ccf(combined$media_pct, combined$arrests, 
    lag.max = 10, na.action = na.pass, 
    main = "Cross-Correlation: media coverage vs. ICE arrests ")
```

A CCF with spikes that all surpass the blue dotted line indicates that each correlation for every time lag is statistically significant. This demonstrates what was previously mathematically calculated. This allows us to visually see that across all the weeks there is a relationship between media coverage on immigration and log of ICE arrests that is statistically significant.

A CCF allows me to determine the similarities between the time series analysis for ICE arrests and Media on immigration, with a time lag placed on ICE arrests. This allows me to see if media coverage was actually influencing the volume of ICE arrests. The first CCF does not contain first-differencing, as a result it is currently including exogenous shocks that may be causing media and ICE arrests to move together. That is why we see each peak surpassing the dotted line, indicting each correlation is statistically significant. Therefore, because the exognous shock of President Trump's second term acted as an omitted variable, increasing both media coverage on immigration and ICE arrests, I completed a CCF documenting the period before Trump's second inauguration, and after his inauguration. This would allow me to determine if the relationship seen across times is consistent in each time period. However, because this is simply looking at correlations, we cannot say there is a causal relationship.

**Dot Plot with Confidence Intervals**

```{r}
library(ggplot2)

cor_plot <-  data.frame(
  lag = c("No Lag", "1-Week Lag", "2-Week Lag", "4-Week Lag", "8-Week Lag"),
  Correlation = round(c(
    no_week_cor$estimate, 
    week1_cor$estimate, 
    week2_cor$estimate, 
    week3_cor$estimate, 
    week4_cor$estimate
  ), 3), 
  P_value = round(c(
    no_week_cor$p.value, 
    week1_cor$p.value, 
    week2_cor$p.value, 
    week3_cor$p.value, 
    week4_cor$p.value
  ), 15), 
  CI_lower = c(
    no_week_cor$conf.int[1], 
    week1_cor$conf.int[1], 
    week2_cor$conf.int[1], 
    week3_cor$conf.int[1], 
    week4_cor$conf.int[1]), 
  CI_upper = c(
    no_week_cor$conf.int[2], 
    week1_cor$conf.int[2], 
    week2_cor$conf.int[2], 
    week3_cor$conf.int[2], 
    week4_cor$conf.int[2])
) |> mutate(
  Significant = case_when(
    P_value < 0.001 ~ "***", 
      P_value < 0.01 ~ "**", 
      P_value < 0.05 ~ "*", 
      TRUE ~ ""
  )
)|> 
  ggplot(aes(x = lag, y = Correlation)) + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), 
               width = 0.2, color = "darkred") + 
  geom_point(size = 4, color = "darkred") + 
  labs(
    title = "Correlation between Media Coverage on Immigration and Log ICE Arrests", 
    x= "Time Lags", 
    y = "Correlation (r)", 
    caption = "Error Bars indicate 95% confidence intervals") + 
      theme_minimal()

cor_plot
```

This dot plot with confidence intervals allows us to see that the highest correlation between media coverage on immigration and log ICE arrests is when there is no lag. This is indicative of the two variables having similar patterns to one another. Yet, for each of the 1, 2, and 4 week lags, there still appears to be a moderate correlation, above .4. This positive correlation indicates that as media coverage on ICE increases, this is likely correlated with an increase in ICE arrests as well. The 8 week lag on the other hand is the lowest correlation, so this time gap between media coverage on immigration and log ICE arrests is likely to be more correlated in shorter intervals of time rather than longer ones.

**CCF Graph Pre-Trump and Post-Trump**

```{r}

pre_trump_ccf <- combined |>
  filter(week < as.Date("2025-01-20"))
post_trump_ccf <- combined |> 
  filter(week >= as.Date("2025-01-20"))

ccf(pre_trump_ccf$media_pct, pre_trump_ccf$arrests, 
    lag.max = 10, na.action = na.pass, 
    main = "Cross-Correlation: media coverage vs.\n ICE arrests Pre-Trump Inauguration ")

```

When looking at the time period before Trump's inauguration, the CCF depicts that at weeks 64,5,6,8,9, and 10, these correlations between media coverage on immigration and log ICE arrests are all statistically significant. However, because all correlations are below the y-axis, then the correlation coefficients are all negative. Which means at these correlations that are statistically significant, and increase in media coverage on immigration is associated with a decrease in log ICE arrests. Likewise, if an increase in ICE log arrests preceded media coverage, then media coverage is likely to be lower. This could be an indicative of the political climate under a liberal president who did not make immigration a forefront of his political platform. Therefore, media coverage on immigration is likely to be more positive during this time, which follows the original hypothesis that positive or neutral media is likely to be associated with a decrease or no change in ICE arrests.

```{r}
ccf(post_trump_ccf$media_pct, post_trump_ccf$arrests, 
    lag.max = 10, na.action = na.pass, 
    main = "Cross-Correlation: media coverage vs.\n ICE arrests  Post-Trump Inauguration")
```

When looking at the time period after Trump's inauguration, the CCF depicts that at weeks 1,2, 5,6,7, and 9,10, these correlations between the media coverage on immigration and log of ICE arrests are statistically significant. However, when analyzing these correlations, they do not appear to be very large, and therefore are unlikely to demonstrate a strong relationship between the two variables, and instead, the variables may move together, but it may not be very consistent. The positive lag value being above the y-axis indicate that when media is produced on immigration, at a later time, log ICE arrests are likely to also trend up. However, there is also a peak at -7, which indicates that when an increase in ICE arrests occurs first, media coverage would be lower. This could be a result of a normalization of ICE arrests that make channels more desensitized to this content, and during the Trump Administration, it would be more unlikely for news channels to want to critique the bureaucratic actions on authority of the president because that makes private news sources a target for the administration.

```{r}

#This correlation is similar to ICE Cream and Shark Attacks. 
lm1 <- lm(log_arrests~media_pct, data = combined)
lm2 <- lm(log_arrests~media_lag1, data = combined)
lm3 <- lm(log_arrests~media_lag4, data = combined)



library(modelsummary)

modelsummary(
  list("No lag" = lm1,
       "1-Week Lag"      = lm2,
       "4-Week Lag"      = lm3),
  stars = TRUE,
  coef_rename = c("media_prop"  = "Media Coverage (Same Week)",
                  "media_lag1"  = "Media Coverage (1-Week Lag)",
                  "media_lag4"  = "Media Coverage (4-Week Lag)"),
  gof_map = c("nobs", "r.squared", "adj.r.squared"),
  title = "National Media Coverage Correlation with Log ICE Arrests"
)
```

This correlation holds extremely high statistical significance; however, it is very concerning to have such high rates because exogenous factors that have not been included may be impacting the outcome and correlation. Instead, this regression is showing that media is having a direct effect on ICE arrests, but there are other factors or even omitted variables that are biasing this relationship and making it appear to have a strong, and significant result, when in reality other variables are explaining more of the variation and relationship between media and ICE arrests. Nevertheless, these regression coefficients indicate:

-   No lag: for every 1 percentage point increase in media coverage on immigration, there would be a 0.097 increase in log arrests. 

-   One week lag: for every 1 percentage point increase in media coverage on immigration, there would be a 0.094 increase in log arrests the following week. 

-   Four week lag,: for every 1 percentage point increase in media coverage on immigration, there should be a 0.090 increase in log arrests in four weeks after.

In addition, I have an low r\^2, which could indicate that if I factor in other variables, including President Trump and his administrative action, this may be able to account for more of the variation between media coverage on immigration and ICE arrests. However, my data sets did not include any other additional factors. Yet, I will control for the Trump administration in the following regression.

```{r}
ggplot(combined, aes(x = week)) +
  geom_line(aes(y = log_arrests, color = "Log Arrests")) +
  geom_line(aes(y = media_prop * 11 + 7, color = "Media Coverage")) +
  scale_y_continuous(
    name = "Log ICE Arrests",
    limits = c(7, 10),
    sec.axis = sec_axis(~(. - 7)/11,
                        name = "Media Coverage (Proportion)")
  ) +
  scale_colour_manual(values = c("Media Coverage" = "darkred", 
                                 "Log Arrests" = "darkblue")) + 
  labs(title = "Media Coverage vs. Log ICE Arrests Over Time",
       x = "Week",
       color = "") 
```

This overlapped time series uses the Log ICE arrests and Media Coverage on immigration in proportion to see if spikes in media coverage on immigration visually matched a spike in ICE arrests, or if these are related with a time lag. During this period, in particular after 2025, both media coverage on immigration and log ICE arrests begin to rise. However, there is a sharp dip before 2025 of ICE arrests, and it had gradually been decreasing from 2023 go 2025. Interestingly enough, the graph depicts that prior to 2025, the volume of media coverage on immigration was gradually increasing, but ICE arrests were falling during this time. This could be due to immigration being a large topic of debate during election season, and candidates would have been campaigning between 2024 to 2025.

Time Split Pre-Trump vs. Post-Trump:

```{r}
# Pre-election period (Oct 2022 - Nov 2024)
combined_pre <- combined %>%
  filter(week < as.Date("2024-11-05"))

# Post-election through inauguration and beyond
combined_post <- combined %>%
  filter(week >= as.Date("2024-11-05"))

# Pre-election
pre_m1 <- lm(log_arrests ~ media_pct, data = combined_pre)
pre_m2 <- lm(log_arrests ~ media_lag1, data = combined_pre)
pre_m3 <- lm(log_arrests ~ media_lag4, data = combined_pre)

# Post-election
post_m1 <- lm(log_arrests ~ media_pct, data = combined_post)
post_m2 <- lm(log_arrests ~ media_lag1, data = combined_post)
post_m3 <- lm(log_arrests ~ media_lag4, data = combined_post)

modelsummary(
  list("No Lag" = pre_m1,
       "1-Week Lag" = pre_m2,
       "4-Week Lag" = pre_m3),
  stars = TRUE,
  coef_rename = c(
    "media_pct"  = "Media Coverage, Same Week (pct)",
    "media_lag1" = "Media Coverage, 1-Week Lag (pct)",
    "media_lag4" = "Media Coverage, 4-Week Lag (pct)"
  ),
  gof_map = c("nobs", "r.squared", "adj.r.squared"),
  title = "Pre-Election: Media Coverage and Log ICE Arrests"
)

modelsummary(
  list("No Lag" = post_m1,
       "1-Week Lag" = post_m2,
       "4-Week Lag" = post_m3),
  stars = TRUE,
  coef_rename = c(
    "media_pct"  = "Media Coverage, Same Week (pct)",
    "media_lag1" = "Media Coverage, 1-Week Lag (pct)",
    "media_lag4" = "Media Coverage, 4-Week Lag (pct)"
  ),
  gof_map = c("nobs", "r.squared", "adj.r.squared"),
  title = "Post-Election: Media Coverage and Log ICE Arrests"
)

```

Next, I held President Trump's inauguration day constant by regressing log ICE arrests on media coverage on immigration before Trump's inauguration, and then after his inauguration. Before his inauguration, media coverage on immigration would negatively impact ICE arrests. The regression coefficients would be:

-   If there was a one percentage point increase in media coverage on immigration, there would be 0.029 decrease in log ICE arrests. However, because this is not statistically significant, we cannot confidently say that this result is real and not simply chance.

-   If there was a one percentage point increase in media coverage on immigration, then after a week, there would be a 0.046 decrease in log ICE arrests. This is statistically significant at the p\<0.05 scale. Therefore, this is more likely to be a real effect and not by chance and indicate that when media coverage on immigration increases, this may be associated with fewer ICE arrests the following week.

-   If there is a one percentage point increase in media coverage on immigration, then after four weeks, there would be a 0.014 decrease in log ICE arrests. However, this result is not statistically significant, so we cannot be confident that this result is real and not simply by chance.

Out of all three regression coefficients, there appears to be the largest change at the one week time lag. This could be due to a more rapid response to ICE bureuacrats who are more pressured to take active response, especially if immigration is becoming increasingly more salient in the news.

In order to better understand the role of media, I looked towards specific polarized news sources in order to determine the effect of charged news sources on politically polarized states. Following the same procedures of before, I filtered the words "immigration," "immigrants," and "border" to represent discussions of immigration. Then I specifically looked at the New York Times and the Guardian to represent more liberal leaning news, and Fox News and the Federalist to represent conservative leaning media sources. I was then able to filter my ICE Arrests data to look at blue and red states specifically, and analyze the relationship between polarized media and the impact it may have on ICE action with various types of polarized states.

National Coverage on Red vs. Blue States

1\) Separate into Blue and Red States

```{r}
blue_states <- c("CALIFORNIA", "NEW YORK", "ILLINOIS", "NEW JERSEY", 
                 "MASSACHUSETTS", "WASHINGTON", "COLORADO", "OREGON",
                 "CONNECTICUT", "MARYLAND", "MINNESOTA", "NEVADA",
                 "NEW MEXICO", "VIRGINIA", "HAWAII", "DELAWARE",
                 "RHODE ISLAND", "VERMONT", "MAINE")

red_states <- c("TEXAS", "FLORIDA", "GEORGIA", "ARIZONA", "TENNESSEE",
                "LOUISIANA", "ALABAMA", "SOUTH CAROLINA", "INDIANA",
                "ARKANSAS", "MISSISSIPPI", "KENTUCKY", "MISSOURI",
                "OKLAHOMA", "NORTH CAROLINA", "UTAH", "IOWA",
                "WEST VIRGINIA", "IDAHO", "WYOMING", "NORTH DAKOTA",
                "SOUTH DAKOTA", "MONTANA", "ALASKA", "OHIO",
                "MICHIGAN", "WISCONSIN", "PENNSYLVANIA", "NEBRASKA",
                "KANSAS")



arrests_split <- arrests_raw %>%
  mutate(state_type = case_when(
    apprehension_state %in% blue_states ~ "blue",
    apprehension_state %in% red_states ~ "red",
    TRUE ~ "other"
  ))

arrests_blue <- arrests_split %>% filter(state_type == "blue")
arrests_red <- arrests_split %>% filter(state_type == "red")




```

Time Series and Regression for Separate State Colors

```{r}
#Red and Blue State Aggregations
blue_weekly <- arrests_split %>%
  filter(state_type == "blue") %>%
  mutate(week = floor_date(as.Date(apprehension_date), "week")) %>%
  group_by(week) %>%
  summarise(arrests = n())

red_weekly <- arrests_split %>%
  filter(state_type == "red") %>%
  mutate(week = floor_date(as.Date(apprehension_date), "week")) %>%
  group_by(week) %>%
  summarise(arrests = n())

#Build combined datasets 
blue_combined <- blue_weekly |>
  inner_join(media_weekly, by = "week") |>
  filter(week != as.Date("2022-09-25")) |>
  arrange(week) |>
  mutate(media_pct = media_prop * 100) |>
  mutate(
    log_arrests = log(arrests),
    media_lag1 = lag(media_pct, 1),
    media_lag4 = lag(media_pct, 4),
    post_trump = ifelse(week >= as.Date("2025-01-20"), 1, 0)
  )

red_combined <- red_weekly |>
  inner_join(media_weekly, by = "week") |>
  filter(week != as.Date("2022-09-25")) |>
  arrange(week) |>
  mutate(media_pct = media_prop * 100) |>
  mutate(
    log_arrests = log(arrests),
    media_lag1 = lag(media_pct, 1),
    media_lag4 = lag(media_pct, 4),
    post_trump = ifelse(week >= as.Date("2025-01-20"), 1, 0)
  )

# Time series plot
bind_rows(
  blue_combined %>% mutate(state_type = "Blue States"),
  red_combined %>% mutate(state_type = "Red States")
) %>%
ggplot(aes(x = week, y = log_arrests, color = state_type)) +
  geom_line() +
  scale_color_manual(values = c("Blue States" = "blue", "Red States" = "red")) +
  geom_vline(xintercept = as.Date("2024-11-05"), linetype = "dashed") +
  geom_vline(xintercept = as.Date("2025-01-20"), linetype = "dashed") +
  annotate("text", x = as.Date("2024-11-05"), y = Inf, label = "2024 election", 
           vjust = 2, hjust = -0.1, size = 3) +
  annotate("text", x = as.Date("2025-01-20"), y = Inf, label = "Trump inauguration", 
           vjust = 4, hjust = -0.1, size = 3) +
  labs(title = "Log ICE Arrests Per Week: Red vs Blue States",
       x = "Week", y = "Log Arrests", color = "") +
  theme_minimal()


```

This time series analysis indicated the volume of log arrests for both Red states and Blue states. Although they both follow similar trends, there is a significant gap between red and blue states, which indicates that red states on average, may be more likely to have high levels of ICE arrest rates than blue states. Therefore, regardless of media, they may always output higher levels of ICE arrests.

# Regressions with Post-Trump Control for Blue and Red States 

```{r}
blue_m1 <- lm(log_arrests ~ media_pct + post_trump, data = blue_combined)
blue_m2 <- lm(log_arrests ~ media_lag1 + post_trump, data = blue_combined)
blue_m3 <- lm(log_arrests ~ media_lag4 + post_trump, data = blue_combined)

red_m1 <- lm(log_arrests ~ media_pct + post_trump, data = red_combined)
red_m2 <- lm(log_arrests ~ media_lag1 + post_trump, data = red_combined)
red_m3 <- lm(log_arrests ~ media_lag4 + post_trump, data = red_combined)

modelsummary(
  list(
    "Blue: No Lag " = blue_m1, 
    "Blue: 1 week Lag " = blue_m2,
    "Blue: 4 week Lag " = blue_m3,
    "Red: No Lag " = red_m1,
    "Red: 1 week  Lag " = red_m2,
    "Red: 4 week Lag " = red_m3
    
  ),
  stars = TRUE, 
  coef_rename = c( 
    "media_pct" = "Media Coverage, Same Week in Percentage", 
    "media_lag1" = "Media Coverage, 1 Week Lag in Percentage", 
    "media_lag4" = "Media Coverage, 4 Week Lag in Percentage", 
    "post_trump" = "Post President Trump's Inauguration"), 
  title = "National Media Coverage on Immigration and Log ICE Arrests: Red vs. Blue States", 
  caption = "Models all control for Trump Inauguration"
)




```

These regressions are specifically looking at the relationship between national media coverage and ICE arrests in either red or blue states. The regression coefficients indicate:

-   If there was a one percentage point increase in national media coverage on immigration, there would be 0.003 increase in log ICE arrests for blue states. However, because this is not statistically significant, we cannot confidently say that this result is real and not simply chance.

-   If there was a one percentage point increase in media coverage on immigration, then after a week, there would be a 0.006 decrease in log ICE arrests for blue states. However, because this is not statistically significant, we cannot confidently say that this result is real and not simply chance. In addition, it is very interesting to see that after a week, national media coverage on immigration no longer is associated with an increase in ICE arrests, but rather a decrease in ICE arrests.

-   If there is a one percentage point increase in media coverage on immigration, then after four weeks, there would be a 0.018 decrease in log ICE arrests for blue states. However, this result is not statistically significant, so we cannot be confident that this result is real and not simply by chance. This continues to follow the pattern of longer lags being associated with decreases in ICE arrests.

-   In addition, because I am controlling for Trump's Inauguration, this covariate coefficient indicates that when President Trump was inaugurated, there would be a increase for log ICE arrests of 0.850 for blue states. This is a very large coefficient and is highly statistically significant which indicates that it is unlikely that we would be receiving this coefficient purely by chance, and there may be a relationship between the two variables.

-   For red states, when President Trump was inaugurated, there appeared to be a 0.939 increase in log ICE arrests. This result was also statistically significant and indicates that this coefficient is not likely to be recieved purely by chance. It is very interesting though that red states have a higher regression coefficient, which indicates that they are more likely to have a greater impact by Trump's inauguration on ICE arrests than red states. This may be a result of red states with previously high rates of arrests were Florida and Texas, and under the Trump administration, these immigration policies were heightened and enhanced which may have led states with already high immigration rates to become even higher.

-   If there was a one percentage point increase in national media coverage on immigration, there would be 0.011 increase in log ICE arrests for red states. However, because this is not statistically significant, we cannot confidently say that this result is real and not simply chance.

-   If there was a one percentage point increase in media coverage on immigration, then after a week, there would be a 0.005 increase in log ICE arrests for red states. However, because this is not statistically significant, we cannot confidently say that this result is real and not simply chance. This result is extremely small and indicates that a change in media coverage is not likely to have any effect on log ICE arrests for red states after one week.

-   If there is a one percentage point increase in media coverage on immigration, then after four weeks, there would be a 0.001 increase in log ICE arrests. However, this result is not statistically significant, so we cannot be confident that this result is real and not simply by chance. This result is extremely small and indicates that a change in media coverage is not likely to have any effect on log ICE arrests for red states after four weeks.

# Blue States Table

```{r}

library(modelsummary)

# Blue States Table
modelsummary(
  list("No Lag"     = blue_m1,
       "1-Week Lag" = blue_m2,
       "4-Week Lag" = blue_m3),
  stars = TRUE,
  coef_rename = c(
    "media_pct"  = "Media Coverage, Same Week (pct)",
    "media_lag1" = "Media Coverage, 1-Week Lag (pct)",
    "media_lag4" = "Media Coverage, 4-Week Lag (pct)",
    "post_trump" = "Post-Trump Inauguration"
  ),
  gof_map = c("nobs", "r.squared", "adj.r.squared"),
  title = "Blue States: National Media Coverage Correlation With Log ICE Arrests"
)

```

This table contains all of the same regression coefficients above, but this is selected to see only the blue state coefficients. However, the interpretations remain the same as stated above. Once again, the Trump Inauguration control is explaining a lot of variation in the data, which may indicate that it is acting as an omitted variable having a large effect on media and ICE arrests.

# Red States Table

```{r}

modelsummary(
  list("No Lag"     = red_m1,
       "1-Week Lag" = red_m2,
       "4-Week Lag" = red_m3),
  stars = TRUE,
  coef_rename = c(
    "media_pct"  = "Media Coverage, Same Week (pct)",
    "media_lag1" = "Media Coverage, 1-Week Lag (pct)",
    "media_lag4" = "Media Coverage, 4-Week Lag (pct)",
    "post_trump" = "Post-Trump Inauguration"
  ),
  gof_map = c("nobs", "r.squared", "adj.r.squared"),
  title = "Red States: National Media Coverage Correlation With Log ICE Arrests"
)

```

This table contains all of the same regression coefficients above, but this is selected to see only the red state coefficients. However, the interpretations remain the same as stated above. Like the blue states, the Trump Inauguration variable is accounting for a large variation in the data which may indicate that it is an omitted variable. In addition, these coefficients are very small and likely indicate that the media has very little effect on log ICE arrests.

Side by Side Trump Control:

```{r}
#Without post_trump control
blue_m1_raw <- lm(log_arrests ~ media_pct, data = blue_combined)
red_m1_raw <- lm(log_arrests ~ media_pct, data = red_combined)

#With post_trump control
blue_m1_controlled <- lm(log_arrests ~ media_pct + post_trump, data = blue_combined)
red_m1_controlled <- lm(log_arrests ~ media_pct + post_trump, data = red_combined)

modelsummary(
  list("Blue (raw)" = blue_m1_raw,
       "Blue (controlled)" = blue_m1_controlled,
       "Red (raw)" = red_m1_raw,
       "Red (controlled)" = red_m1_controlled),
  stars = TRUE,
  coef_rename = c(
    "media_pct" = "Media Coverage, Same Week (pct)",
    "post_trump" = "Post-Trump Inauguration"
  ),
  gof_map = c("nobs", "r.squared", "adj.r.squared"),
  title = "Red vs Blue States: Effect of Controlling for Trump Inauguration"
)

```

This table allows us to see how much the coefficients for media coverage change when the control of Post-Trump Inauguration is included. In particular, for both blue and red states, media coverage went from being a statistically significant coefficient to not, and losing much of its explanatory power. This indicates that media may not have as much explanatory power as originally thought.

Liberal Media Impact on Blue States:

```{r}
liberal_weekly <- liberal_media %>%
  mutate(week = floor_date(as.Date(date), "week")) %>%
  group_by(week) %>%
  summarise(
    liberal_count = sum(count),
    liberal_total = sum(total_count),
    media_prop = liberal_count / liberal_total
  )

blue_combined_sourced <- blue_weekly |>
  inner_join(liberal_weekly, by = "week") |>
  filter(week != as.Date("2022-09-25")) |>
  arrange(week) |>
  mutate(media_pct = media_prop * 100) |>
  mutate(
    log_arrests = log(arrests),
    media_lag1 = lag(media_pct, 1),
    media_lag4 = lag(media_pct, 4),
    post_trump = ifelse(week >= as.Date("2025-01-20"), 1, 0)
  )

blue_src_m1 <- lm(log_arrests ~ media_pct + post_trump, data = blue_combined_sourced)
blue_src_m2 <- lm(log_arrests ~ media_lag1 + post_trump, data = blue_combined_sourced)
blue_src_m3 <- lm(log_arrests ~ media_lag4 + post_trump, data = blue_combined_sourced)

modelsummary(
  list("No Lag"     = blue_src_m1,
       "1-Week Lag" = blue_src_m2,
       "4-Week Lag" = blue_src_m3),
  stars = TRUE,
  coef_rename = c(
    "media_pct"  = "Liberal Media Coverage, Same Week (pct)",
    "media_lag1" = "Liberal Media Coverage, 1-Week Lag (pct)",
    "media_lag4" = "Liberal Media Coverage, 4-Week Lag (pct)",
    "post_trump" = "Post-Trump Inauguration"
  ),
  gof_map = c("nobs", "r.squared", "adj.r.squared"),
  title = "Blue States: Liberal Media Coverage and Log ICE Arrests"
)




```

When looking specifically at the impact of liberal media coverage on blue states, each coefficient indicates:

-   If there was a one percentage point increase in liberal national media coverage on immigration, there would be 0.032 increase in log ICE arrests for blue states. This is statistically significant at the p\< 0.01 level, and this indicates that it is very unlikely that these results are due to chance and not a relationship between the two variables.

-   If there was a one percentage point increase in liberal national media coverage on immigration, then after a week, there would be a 0.028 increase in log ICE arrests for blue states. This is also statistically significant at the p\< 0.01 level, and this indicates that it is very unlikely that these results are due to chance.

-   If there is a one percentage point increase in liberal national media coverage on immigration, then after four weeks, there would be a 0.024 increase in log ICE arrests for blue states. This is statistically significant at the p\< 0.05 level, and this indicates that it is very unlikely that these results are due to chance.

-   In addition, because I am controlling for Trump's Inauguration, this covariate coefficient indicates that when President Trump was inaugurated, there would be a increase for log ICE arrests of 1.122 for blue states. This is a very large coefficient and is highly statistically significant at the p\< 0.001 which indicates that it is unlikely that we would be receiving this coefficient purely by chance, and there may be a relationship between the two variables.

With all three coefficients being statistically significant, this indicates that there is likely a relationship between liberal national media coverage on immigration and ICE arrests in blue states. Nevertheless, these results are not very large, but they are very likely.

Just looking at the comparison of the correlation Pre and Post-Trump's Inaugaration:

```{r}
blue_src_pre <- blue_combined_sourced %>%
  filter(week < as.Date("2025-01-20"))

blue_src_post <- blue_combined_sourced %>%
  filter(week >= as.Date("2025-01-20"))

blue_pre_m1 <- lm(log_arrests ~ media_pct, data = blue_src_pre)
blue_pre_m2 <- lm(log_arrests ~ media_lag1, data = blue_src_pre)
blue_pre_m3 <- lm(log_arrests ~ media_lag4, data = blue_src_pre)

# Blue post-inauguration regressions
blue_post_m1 <- lm(log_arrests ~ media_pct, data = blue_src_post)
blue_post_m2 <- lm(log_arrests ~ media_lag1, data = blue_src_post)
blue_post_m3 <- lm(log_arrests ~ media_lag4, data = blue_src_post)

modelsummary(
  list("Pre-Inauguration, No Lag"     = blue_pre_m1,
       "Pre-Inauguration, 1-Week Lag" = blue_pre_m2,
       "Pre-Inauguration, 4-Week Lag" = blue_pre_m3,
       "Post-Inauguration, No Lag"    = blue_post_m1,
       "Post-Inauguration, 1-Week Lag"= blue_post_m2,
       "Post-Inauguration, 4-Week Lag"= blue_post_m3),
  stars = TRUE,
  coef_rename = c(
    "media_pct"  = "Liberal Media Coverage, Same Week (pct)",
    "media_lag1" = "Liberal Media Coverage, 1-Week Lag (pct)",
    "media_lag4" = "Liberal Media Coverage, 4-Week Lag (pct)"
  ),
  gof_map = c("nobs", "r.squared", "adj.r.squared"),
  title = "Blue States: Liberal Media Coverage and Log ICE Arrests, Pre vs Post Inauguration"
)


```

Conservative Media Impact on Red States:

```{r}
conservative_weekly <- conservative_media %>%
  mutate(week = floor_date(as.Date(date), "week")) %>%
  group_by(week) %>%
  summarise(
    conservative_count = sum(count),
    conservative_total = sum(total_count),
    media_prop = conservative_count / conservative_total
  )

red_combined_sourced <- red_weekly |>
  inner_join(conservative_weekly, by = "week") |>
  filter(week != as.Date("2022-09-25")) |>
  arrange(week) |>
  mutate(media_pct = media_prop * 100) |>
  mutate(
    log_arrests = log(arrests),
    media_lag1 = lag(media_pct, 1),
    media_lag4 = lag(media_pct, 4),
    post_trump = ifelse(week >= as.Date("2025-01-20"), 1, 0)
  )

red_src_m1 <- lm(log_arrests ~ media_pct + post_trump, data = red_combined_sourced)
red_src_m2 <- lm(log_arrests ~ media_lag1 + post_trump, data = red_combined_sourced)
red_src_m3 <- lm(log_arrests ~ media_lag4 + post_trump, data = red_combined_sourced)

modelsummary(
  list("No Lag"     = red_src_m1,
       "1-Week Lag" = red_src_m2,
       "4-Week Lag" = red_src_m3),
  stars = TRUE,
  coef_rename = c(
    "media_pct"  = "Conservative Media Coverage, Same Week (pct)",
    "media_lag1" = "Conservative Media Coverage, 1-Week Lag (pct)",
    "media_lag4" = "Conservative Media Coverage, 4-Week Lag (pct)",
    "post_trump" = "Post-Trump Inauguration"
  ),
  gof_map = c("nobs", "r.squared", "adj.r.squared"),
  title = "Red States: Conservative Media Coverage and Log ICE Arrests"
)

```

-   This is specifically analyzing the relationship between conservative media coverage and Log ICE arrests in red states, holding the Trump Inauguration constant. The coefficients indicate:

-   If there was a one percentage point increase in conservative national media coverage on immigration, there would be 0.001 increase in log ICE arrests for red states. This is not statistically significant, and indicative of this association being posisbly due to chance.

-   If there was a one percentage point increase in conservative national media coverage on immigration, then after a week, there would be a 0.002 decrease in log ICE arrests for red states. This is also not statistically significant, and indicative of this association being possibly due to chance.

-   If there is a one percentage point increase in conservative national media coverage on immigration, then after four weeks, there would be a 0.004 decrease in log ICE arrests for red states. Last but not least, this was also is not statistically significant, and indicative of this association being possibly due to chance.

-   In addition, because I am controlling for Trump's Inauguration, this covariate coefficient indicates that when President Trump was inaugurated, there would be a increase for log ICE arrests of 0.946 for red states. This is a very large coefficient and is highly statistically significant at the p\< 0.001 which indicates that it is unlikely that we would be receiving this coefficient purely by chance, and there may be a relationship between the two variables. In addition, this is much a larger coefficient than any of our media variables on the log ICE arrests which indicates that Trump's inauguration may have large explanatory power to reveal how log ICE arrests are changing over time.

Red Media Split Pre and Post-Trump

```{r}
red_src_pre <- red_combined_sourced %>%
  filter(week < as.Date("2025-01-20"))

red_src_post <- red_combined_sourced %>%
  filter(week >= as.Date("2025-01-20"))

red_pre_m1 <- lm(log_arrests ~ media_pct, data = red_src_pre)
red_pre_m2 <- lm(log_arrests ~ media_lag1, data = red_src_pre)
red_pre_m3 <- lm(log_arrests ~ media_lag4, data = red_src_pre)


red_post_m1 <- lm(log_arrests ~ media_pct, data = red_src_post)
red_post_m2 <- lm(log_arrests ~ media_lag1, data = red_src_post)
red_post_m3 <- lm(log_arrests ~ media_lag4, data = red_src_post)


modelsummary(
  list("Pre-Inauguration, No Lag"     = red_pre_m1,
       "Pre-Inauguration, 1-Week Lag" = red_pre_m2,
       "Pre-Inauguration, 4-Week Lag" = red_pre_m3,
       "Post-Inauguration, No Lag"    = red_post_m1,
       "Post-Inauguration, 1-Week Lag"= red_post_m2,
       "Post-Inauguration, 4-Week Lag"= red_post_m3),
  stars = TRUE,
  coef_rename = c(
    "media_pct"  = "Conservative Media Coverage, Same Week (pct)",
    "media_lag1" = "Conservative Media Coverage, 1-Week Lag (pct)",
    "media_lag4" = "Conservative Media Coverage, 4-Week Lag (pct)"
  ),
  gof_map = c("nobs", "r.squared", "adj.r.squared"),
  title = "Red States: Conservative Media Coverage and Log ICE Arrests, Pre vs Post Inauguration"
)


```

Limitations:

Although the time series offers high external validity due to its analysis on actual media coverage and its relationship to ICE arrests that have been documented, it lacks strong internal validity. Because we are examining the correlation between the two variables, we cannot claim that this is a causal relationship. In fact, even if sentiment coverage and ICE arrests move together on the time series, we do not know if ICE agents are watching the cable news channels highlighted through Television News Archive. Without this direct measurement, we cannot determine if ICE bureaucrats are actually responding to consumed media. In addition, because we are conducting national level testing, it is difficult to analyze whether specific bureaucratic agents are deliberately changing their behavior based on national media. However, it is difficult to conduct state level testing due to the lack of captions and full segments from local state channels. Although the FCC requires channels to broadcast captions while live streaming, it does not require them to publish them after. It would be impossible to get all captions and segments from local news channels to conduct this analysis. 

Confounding variables may actually be driving media coverage and ICE arrests independently. For example, political events, such as Trump’s campaign and election in 2024, border crossing surges, and any policies passed by the congressional or executive branch, including Trump’s expansion of the DHS, may have both influenced arrests and brought additional coverage to the topic. In addition, it may be plausible that ICE arrests are actually shaping media coverage, and the time series, as a result, may reveal a reverse causal relationship. In fact, specific highly polarized or viral media may reach larger audiences including ICE bureaucrats, but this study treats all negative segments equally regardless of intensity, and applies the same categorization to positive segments. As a result, it is unlikely I can analyze the nuance and the volatility of certain immigration stories over others. 

 Many researchers have studied the influence of the media on public opinion, but few have analyzed the media as a direct signal to bureaucratic agencies. This paper examined the relationship between media coverage of immigration on ICE arrests nationwide. My goal with conducting this research is to evaluate whether the media distorts the opinions of agents who are supposed to be apolitical. Understanding the direct influence of the media on agencies could influence the way that policy makers regulate media because of its potential to influence policy that gets implemented.

Reference List 

Bordonada, M. Royo. 2022. “Corporate capture of public health by food industries in Europe.” European Journal of Public Health 32, (3). https://doi.org/10.1093/eurpub/ckac129.091. 

Chavez, Leo R. 2013. The Latino Threat: Constructing Immigrants, Citizens, and the Nation, Second Edition. Stanford University Press. 

Cohen, Bernard C. 1963. The Press and Foreign Policy. Princeton University Press. 

Cohen, Stanley. 2002. Folk Devils and Moral Panics: The creation of the Mods and Rockers, Third Edition. Routledge. 

DeMarzo, Peter M., Dimitri Vayanos, and Jeffrey Zwiebel. 2003. “Persuasion Bias, Social Influence, and Unidimensional Opinions.” The Quarterly Journal of Economics 118, (3). https://doi.org/10.1162/00335530360698469.

Domhoff, William G., Ronald W. Cox, John L. Campbell, Richard W. Lachmann, Clarence Y.H. Lo, Beth Mintz, Joseph G. Peschek, Robert J.S. Ross, Daniel J. Schneider, Michael Schwartz, Kathleen C. Schwartzman, and Judith Stephan-Norris. 2017. Studying the Power Elite: Fifty Years of Who Rules America? Routledge Taylor and Francis Group. 

“Deportation Data Project.” Deportation Data Project, deportationdata.org/. 

Gilens, Martin and Craig Hertzman. 2003. “Corporate Ownership and News Bias: Newspaper Coverage of the 1996 Telecommunications Act.” Journal of Politics 62, (2): 369-386. https://doi.org/10.1111/0022-3816.00017. 

Huber, John D. and Michael M. Ting. 2015. “Civil Service and Patronage in Democracies.” Working paper, Department of Political Science, Columbia University. polisci.columbia.edu/sites/polisci.columbia.edu/files/content/pdfs/Publications/Huber/Working%20Papers/huber%20ting%20personnel.

“ICE Administrative Arrests.” Deportation Data Project, Oct. 2022- Mar. 2026, ice-arrests.apps.deportationdata.org. 

Iyengar, Shanto and Donald R. Kinder. 1987. News that Matters: Television and American Opinion. University of Chicago Press. 

McCombs, Maxwell E., and Donald L. Shaw. 1972. “The Agenda-Setting Functions of Mass Media.” The Public Opinion Quarterly 36, (2): 176-87. http://jstor.org/stable/2747787.

Michael P. Vandenbergh, J. M. Gilligan, and H. Feurman. 2020. The New Revolving Door, 70 Case Western Reserve Law Review. 1121. https://doi.org/10.2139/SSRN.3570889. 

Newport, Frank. 2018 “Immigration Jumps as Top Problem, Still Trails Government.” GALLUP, February 15. news.gallup.com/poll/227021/immigration-jumps-top-problem-trails-government.aspx.

Saad, Lydia. 2014. “One in Six Say Immigration Most Important U.S. Problem.” GALLUP, July 16. news.gallup.com/poll/173306/one-six-say-immigration-important-problem.aspx.

“Television News Archive.” Internet Archive, archive.org/details/tv. 

“Unauthorized Immigrant Population Profiles.” Migration Policy Institute, www.migrationpolicy.org/programs/us-immigration-policy-program-data-hub/unauthorized-immigrant-population-profiles. 

Walker, Edward T. 2009. “Privatizing Participation: Civic Change and the Organizational Dynamics of Grassroots Lobbying Firms.” American Sociological Review 74, (1). https://doi.org/10.1177/000312240907400105.

Weber, Max. 1946. “Bureaucracy.” In From Max Weber: Essays in Sociology, edited by H.H. Gerth and C. Wright Mills, Oxford University Press. https://doi.org/10.4324/9780203452196.

Yackee, Jason Webb and Susan Webb Yackee. 2006. “A Bias Towards Business? Assessing Interest Group Influence on the U.S. Bureaucracy.” The Journal of Politics 68, (1): 128-139. https://doi.org/10.1111/j.1468-2508.2006.00375.
